---
title: "hw4 Time series"
author: '106070020'
date: "2021年4月10日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3.10(a)

```{r}
library(astsa)
(regr<-ar.ols(cmort, order=2,demean=F, intercept=T))
regr$asy.se.coef
```

***
>Our estimates are Phi0=11.45 se(2.394), Phi1=0.43 (se=.04), Phi2=0.44 (se=.04), and sigma^2=32.32. Est. model --> Xt = 11.45 + 0.43*X(t-1) + 0.44*X(t-2)


## 3.10(b)
```{r}
# Find 4 week forcasts and 95% PI's for model (a)
cmort.pred <-predict(regr,n.ahead=4)
# list the 95% PI for each increase
upper <-c(cmort.pred$pred+2*cmort.pred$se)
lower <-c(cmort.pred$pred-2*cmort.pred$se)
list(m1=c(lower[1],upper[1]),m2=c(lower[2],upper[2]),
     m3=c(lower[3],upper[3]),m4=c(lower[4],upper[4]))
```

## 3.10(c)
```{r message=F}
bound<-function(mean,sd,lambda){
  c(mean-3*sd*sqrt(lambda/(2-lambda)),mean+3*sd*sqrt(lambda/(2-lambda)))
}
library(zoo)
t<-index(cmort)
library(qcc)

l0.8<-ewmaSmooth(t,cmort,lambda=0.8)
l0.6<-ewmaSmooth(t,cmort,lambda=0.6)
l0.5<-ewmaSmooth(t,cmort,lambda=0.5)
l0.2<-ewmaSmooth(t,cmort,lambda=0.2)

par(mfrow=c(2,2))
{plot(l0.8,type="l", main="EWMA with lambda=0.8", ylim=c(60,130))
abline(h=bound(mean(cmort),sd(cmort),0.8)[1], col="red")
abline(h=bound(mean(cmort),sd(cmort),0.8)[2], col="red")
abline(v=150,col="blue",lty="dotted")}
{plot(l0.6,type="l", main="EWMA with lambda=0.6", ylim=c(60,130))
abline(h=bound(mean(cmort),sd(cmort),0.6)[1], col="red")
abline(h=bound(mean(cmort),sd(cmort),0.6)[2], col="red")
abline(v=150,col="blue",lty="dotted")}
{plot(l0.5,type="l",main="EWMA with lambda=0.5", ylim=c(60,130))
abline(h=bound(mean(cmort),sd(cmort),0.5)[1], col="red")
abline(h=bound(mean(cmort),sd(cmort),0.5)[2], col="red")
abline(v=150,col="blue",lty="dotted")}
{plot(l0.2,type="l", main="EWMA with lambda=0.2", ylim=c(60,130))
abline(h=bound(mean(cmort),sd(cmort),0.2)[1], col="red")
abline(h=bound(mean(cmort),sd(cmort),0.2)[2], col="red")
abline(v=150,col="blue",lty="dotted")}
sse<-c(sum((l0.8$y-cmort)^2),
       sum((l0.6$y-cmort)^2),
       sum((l0.5$y-cmort)^2),
       sum((l0.2$y-cmort)^2))
sse<-as.data.frame(sse)
rownames(sse)<-c('lambda=0.8','lambda=0.6', 'lambda=0.5', 'lambda=0.2')
sse

```

***
> Choose lambda=0.8, as the process is in control because most of the values are in the upper and lower bound, which are defined by control limits for EWMA. Also, by setting lambda=0.8 we can get the minimum sse. 

## 3.10(d)
```{r}
arpred<-predict(regr,n.ahead=2)
arpred$pred
m2<-c(lower[2],upper[2])
m2
l0.8$y[490] #start:1979.404
l0.8$y[491] #end:1979.423
```

***
> We can see the result of 1-step-ahead prediction of ewma model when lambda=0.8 are both in the 95% prediction interval in the AR(2) model. 




















